---
title: "Rideau Canal"
output: html_document
---

```{r}
library(lubridate)
library(tidyverse)
```

Import the dates on which the Rideau Canal Skateway was opened. This input file was originally copied from [Wikipedia](https://en.wikipedia.org/w/index.php?title=Rideau_Canal&oldid=995637706#Skateway "Wikipedia article: Rideau Canal, section Skateway, revision 995637706").

```{r Import dates}
# Import data
dates <- read.csv("rideau_canal_dates.csv", stringsAsFactors = F, header = T, row.names = 1)

dates$date_fmt <- as.Date(dates$date, format = "%B %d, %Y")

dates$month <- as.numeric(format(dates$date_fmt, "%m"))
dates$day <- as.numeric(format(dates$date_fmt, "%d"))
```

Find the earliest and latest opening dates.
```{r Find the earliest and latest opening dates}

# Get minimum and maximum dates
# It's sorta tricky since a season spans two years (e.g. Dec 2011 J- an 2012)
# For the minimum day look at the earliest year-day before June; for max after June
min_day <- min(yday(dates[month(dates$date_fmt) > 6,]$date_fmt))
min_day <- subset(dates, yday(date_fmt)==min_day)[1,] # In a tie just take the first one

max_day <- max(yday(dates[month(dates$date_fmt) < 6,]$date_fmt))
max_day <- subset(dates, yday(date_fmt)==max_day)[1,] # In a tie just take the first one

cat("Earliest and latest opening days:\n")
cat(min_day$date)
cat("\n")
cat(max_day$date)

```

Add some other useful variables. Add a season variable and a variable counting the days from December 1. Season is defined as the first year (i.e. the 1970-71 season is "1971").
```{r Add season}
# start the axis from Dec 1?
# january + 30 gives days from dec 1
# february + 31 + 30 gives days from dec 1
# dec just use days - 1
dates$from_dec1 <- with(dates,
                        ifelse(month==1, day + 30,
                               ifelse(month==2, day + 31 + 30,
                               day - 1)))

# add season (1970-1971 is the 1970 season)
dates$season <- 2019:(2019-nrow(dates) + 1)
```

Run a linear regression of days after December 1 versus opening day across the entire data set. Obviously this is only an approximation (since eventually it will be possible to not have an opening day).
```{r Regression model and diagnostics}
# regression model
reg_model <- lm(season ~ from_dec1, data = dates)
summary(reg_model)
plot(reg_model)
```

```{r}
# on average, each year the skating season has opened 0.72 days later
# (give or take 0.34 either way)

# how many times did it open in Jan?
# how often has that happened recently?

nrow(dates[dates$month==1,])
nrow(dates[dates$month==12 & dates$season>=2000,]) # seasons since 2000-2001 starting in Jan
nrow(dates[dates$season>=2000,])
nrow(dates[dates$month==12 & 1980 <= dates$season & dates$season < 2000,]) # seasons before 2000-2001 starting in Jan


```

```{r Plot the dates}
ggplot(dates) +
  geom_point(aes(x=season, y=from_dec1)) +
  geom_hline(yintercept=31, linetype = "dotted") +
  geom_smooth(aes(x=season, y=from_dec1),
              method = "lm", formula = y ~ x, se = F, colour="red", linetype="dashed") +
  annotate("text", 1968, 32.4, label="Jan 1", size=3) +
  ylab("Days after Dec 1") + xlab("Season") +
  ggtitle("How late in the season does the Rideau Canal Skateway open?")

ggplot(dates) +
  geom_point(aes(x=season, y=from_dec1)) +
  geom_hline(yintercept=31, linetype = "dotted") +
  annotate("text", 1968, 32.4, label="Jan 1", size=3) +
  ylab("Days after Dec 1") + xlab("Season") +
  ggtitle("How late in the season does the Rideau Canal Skateway open?")
```