---
title: "Rideau Canal"
output: html_document
---

```{r}
library(lubridate)
library(tidyverse)
```

Import the dates on which the Rideau Canal Skateway was opened. This input file was originally copied from [Wikipedia](https://en.wikipedia.org/w/index.php?title=Rideau_Canal&oldid=995637706#Skateway "Wikipedia article: Rideau Canal, section Skateway, revision 995637706").

```{r Import dates}
# Import data
dates <- read.csv("rideau_canal_dates.csv", stringsAsFactors = F, header = T)#, row.names = 1)

dates$opening_date_fmt <- as.Date(dates$opening_date, format = "%B %d, %Y")

dates$month <- as.numeric(format(dates$opening_date_fmt, "%m"))
dates$day <- as.numeric(format(dates$opening_date_fmt, "%d"))
```

Find the earliest and latest opening dates.
```{r Find the earliest and latest opening dates}

# Get minimum and maximum dates
# It's sorta tricky since a season spans two years (e.g. Dec 2011 J- an 2012)
# For the minimum day look at the earliest year-day before June; for max after June
min_day <- min(yday(dates[month(dates$opening_date_fmt) > 6,]$opening_date_fmt))
min_day <- subset(dates, yday(opening_date_fmt)==min_day)[1,] # In a tie just take the first one

max_day <- max(yday(dates[month(dates$opening_date_fmt) < 6,]$opening_date_fmt))
max_day <- subset(dates, yday(opening_date_fmt)==max_day)[1,] # In a tie just take the first one

cat("Earliest and latest opening days:\n")
cat(min_day$opening_date)
cat("\n")
cat(max_day$opening_date)

```

Add some other useful variables. Add a season variable and a variable counting the days from December 1. Season is defined as the first year (i.e. the 1970-71 season is "1971").
```{r Add season}
# start the axis from Dec 1?
# january + 30 gives days from dec 1
# february + 31 + 30 gives days from dec 1
# dec just use days - 1
dates$from_dec1 <- with(dates,
                        ifelse(month==1, day + 30,
                               ifelse(month==2, day + 31 + 30,
                               day - 1)))

# add season (1970-1971 is the 1970 season)
# if the season opens in December, use that year
# Otherwise use the opening date's year minus one
dates$season <- with(dates, ifelse(
  month(opening_date_fmt)==12, year(opening_date_fmt), year(opening_date_fmt)  -1
))
#dates$season <- rev(1970:(nrow(dates) - 1 + 1970))
```

Run a linear regression of days after December 1 versus opening day across the entire data set. Obviously this is only an approximation (since eventually it will be possible to not have an opening day).
```{r Regression model and diagnostics}
# regression model
reg_model <- lm(season ~ from_dec1, data = dates)
summary(reg_model)
plot(reg_model)
```

```{r}
# on average, each year the skating season has opened 0.72 days later
# (give or take 0.34 either way)

# how many times did it open in Jan?
# how often has that happened recently?

nrow(dates[dates$month==1,])
nrow(dates[dates$month==12 & dates$season>=2000,]) # seasons since 2000-2001 starting in Jan
nrow(dates[dates$season>=2000,])
nrow(dates[dates$month==12 & 1980 <= dates$season & dates$season < 2000,]) # seasons before 2000-2001 starting in Jan


```

```{r Plot the dates}
ggplot(dates) +
  geom_point(aes(x=season, y=from_dec1)) +
  geom_hline(yintercept=31, linetype = "dotted") +
  geom_smooth(aes(x=season, y=from_dec1),
              method = "lm", formula = y ~ x, se = F, colour="red", linetype="dashed") +
  annotate("text", 1968, 32.4, label="Jan 1", size=3) +
  ylab("Days after Dec 1") + xlab("Season") +
  ggtitle("How late in the season does the Rideau Canal Skateway open?")

ggplot(dates) +
  geom_point(aes(x=season, y=from_dec1)) +
  geom_hline(yintercept=31, linetype = "dotted") +
  annotate("text", 1968, 32.4, label="Jan 1", size=3) +
  ylab("Days after Dec 1") + xlab("Season") +
  ggtitle("How late in the season does the Rideau Canal Skateway open?")
```
```{r Try to improve the plots}
ggplot(dates) +
  geom_point(aes(x=season, y=from_dec1)) +
  geom_hline(yintercept=31, linetype = "dotted") +
  geom_hline(yintercept=62, linetype = "dotted") +
  geom_smooth(aes(x=season, y=from_dec1),
              method = "lm", formula = y ~ x, se = F, colour="red", linetype="dashed") +
  annotate("text", 1968, 32.4, label="Jan 1", size=3) +
  annotate("text", 1968, 63.4, label="Feb 1", size=3) +
  ylab("Days after Dec 1") + xlab("Season") +
  ggtitle("How late in the season does the Rideau Canal Skateway open?")

ggplot(dates) +
  geom_point(aes(x=season, y=from_dec1)) +
  geom_hline(yintercept=31, linetype = "dotted") +
  geom_hline(yintercept=62, linetype = "dotted") +
  #geom_smooth(aes(x=season, y=from_dec1),
  #            method = "lm", formula = y ~ x, se = F, colour="red", linetype="dashed") +
  annotate("text", 1968, 32.4, label="Jan 1", size=3) +
  annotate("text", 1968, 63.4, label="Feb 1", size=3) +
  ylab("Days after Dec 1") + xlab("Season") +
  ggtitle("How late in the season does the Rideau Canal Skateway open?") #+
  #geom_point(aes(x=2022,y=64),colour="red") # assume Feb 3 opening
```

```{r Trying to make floating bar plots}
# Floating bar plots:
# A bunch of vertical bars where the bottom starts at the opening date and the top goes to the closing date
dates_floating <- read.csv("rideau_canal_dates.csv", stringsAsFactors = F, header = T)#, row.names = 1)

# Format the data
dates_floating$opening_date_formatted <- as.Date(dates_floating$opening_date, format = "%B %d, %Y")
dates_floating$closing_date_formatted <- as.Date(dates_floating$closing_date_fmt, format = "%Y-%m-%d")

# Drop some variables so we define them solely by date
# Also rename them, this dataset became a bit of a hodge podge over the years....
dates_floating <- dates_floating[,c("opening_date_formatted","closing_date_formatted")]

# Add the individual day and month
dates_floating$opening_date_month <- month(dates_floating$opening_date_formatted)
dates_floating$opening_date_day <- day(dates_floating$opening_date_formatted)
dates_floating$closing_date_month <- month(dates_floating$closing_date_formatted)
dates_floating$closing_date_day <- day(dates_floating$closing_date_formatted)

# Set the dates relative to December 1st of each season
dates_floating$opening_from_dec1 <- with(dates_floating,
                                         ifelse(opening_date_month==1, opening_date_day + 30,
                                                ifelse(opening_date_month==2, opening_date_day + 31 + 30,
                                                       opening_date_day - 1)))

# Get the date between the opening and closing
dates_floating$closing_from_dec1 <- with(dates_floating,
                                         as.integer(opening_from_dec1 + (closing_date_formatted - opening_date_formatted)))

# Middle of season marker
dates_floating$mid <- with(dates_floating, as.integer(opening_from_dec1 + (closing_from_dec1 - opening_from_dec1)/2))

# add season (1970-1971 is the 1970 season)
dates_floating$season <- with(dates_floating, ifelse(
  month(opening_date_formatted)==12, year(opening_date_formatted), year(opening_date_formatted)  -1
))

# Add the 2022-23 season (empty)
dates_floating_missing <- dates_floating[0,c("opening_date_formatted","closing_date_formatted")]
dates_floating_missing[1,] <- list(as.Date("2022-12-01", format = "%Y-%m-%d"),
                                as.Date("2023-03-31", format = "%Y-%m-%d"))
dates_floating_missing$opening_from_dec1 <- 1
dates_floating_missing$closing_from_dec1 <- 120
dates_floating_missing$mid <- 60
dates_floating_missing$season <- 2022

# bar width
w <- 0.9

p <- ggplot() +
  geom_crossbar(data = dates_floating,
                aes(x = season, y = mid, ymin = opening_from_dec1, ymax = closing_from_dec1),
                width = w, colour = "blue", fill = "blue", fatten = NULL) +
  geom_crossbar(data = dates_floating_missing,
                aes(x = season, y = mid, ymin = opening_from_dec1, ymax = closing_from_dec1),
                #width = 1, colour = alpha("red", 0.3), fill = "red", alpha=0.3, fatten = NULL) +
                width = w, colour = NA, fill = "red", alpha=0.3, fatten = NULL) +
  xlab("Season") +
  ylab("Season length (starting December 1)") +
  coord_cartesian(ylim = c(0,120))

#+
#  coord_flip() +
#  xlab("Item") + 
#  ylab("Value")

#dates_floating$month <- as.numeric(format(dates_floating$date_fmt, "%m"))
#dates_floating$day <- as.numeric(format(dates_floating$date_fmt, "%d"))
```






